---
title: "How the Bronzong simulator works"
author: "Kevin Z. Lin"
date: "2026-08-29"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Overview

This is a walkthrough of the simulator as it stands, built to be argued with.
The policy it runs — `R/decision_claude.R` — is a **first draft translation of
`docs/03_decision_tree.md`**, written so the tree and the code can be read side
by side. It will make choices you disagree with. Those disagreements are the
output this document is for.

**The question the whole thing answers:** over many simulated games, how often
can the player attack with **Evolution Jammer on or before their own turn 2**?
One fixed bar, the same going first and going second (ADR 0004), reported
separately for the two coin-flip branches and never pooled (ADR 0002).

| Stage | What it does | Functions |
|---|---|---|
| Setup | Shuffle, mulligan, place the Active, set prizes | `setup_game()`, `policy_placement()` |
| Turns 1–2 | Draw, then play the decision tree | `policy_turn()`, `play_replicate()` |
| Measure | One replicate → hit / miss, turn achieved, diagnosis | `summarise_replicate()` |
| Aggregate | Many replicates → one rate per cell | `summarise_run()` |
| Diagnose | A stratified sample of readable traces | `write_trace_file()` |

Two things the simulator is careful about, because both are easy to get silently
wrong:

- **The policy may not see hidden information** (ADR 0003). It reads the hand,
  the board and the discard, plus a belief state that only learns what a real
  player would learn. A search that fails because the card was prized is
  information the player *earns*.
- **Mulligans are never a miss** (ADR 0005). A hand with no Basic is redrawn and
  the redraw is played; the cost is reported beside the rate, never folded into
  it.

```{r libraries, message=FALSE}
for(one_file in list.files("../R", pattern = "[.]R$", full.names = TRUE)){
  source(one_file)
}

card_df <- build_card_database()
decklist <- read_decklist("../decklists/decklist2.txt", card_df)
```

# The deck

`decklist2`, chosen because it runs **both** Salvatore and Latias ex — so the
turn-1 kill line and the free-retreat rung of the §4.3 ladder both get
exercised.

```{r show_decklist}
count_vec <- decklist$count_vec
data.frame(count = as.integer(count_vec),
           card = lookup_card(card_df, names(count_vec))$name,
           row.names = NULL)
```

The four sub-goals that must all hold at the attack step, from §1 of the
decision tree:

| | Sub-goal | Satisfied by |
|---|---|---|
| **A** | A Bronzor is in play | setup, Poké Pad, Ultra Ball, Brock's Scouting, benching one |
| **B** | Bronzong is on it | Hilda, Salvatore, or drawn and evolved |
| **C** | Bronzong is **Active** | led at setup, free retreat, Switch, Surfer, Run Around, Cursed Blast |
| **D** | A `[P]` source is attached | basic Psychic or Telepathic Psychic Energy |

# One game, played slowly

The most useful part of this document. A single seed, going second, played turn
by turn — every decision the policy takes is logged as it happens.

```{r one_game_setup}
seed_number <- 8L

pair <- setup_game(decklist, card_df,
                   bool_going_first = FALSE,
                   placement_fn = policy_placement,
                   seed_number = seed_number)

print(paste0("opening hand: ",
             paste0(lookup_card(card_df, pair$state$hand_vec)$name,
                    collapse = ", ")))
print(paste0("led: ", lookup_card(card_df, top_card(pair$state$active))$name))
print(paste0("bench: ", length(pair$state$bench_list), " Pokemon"))
```

Note the empty Bench. §3 says to place the Active and stop: a benched Basic can
never be un-benched, and the slots are wanted for Latias ex.

## Turn 1

```{r one_game_turn1}
pair$state <- begin_turn(pair$state)
pair <- draw_to_hand(pair, num_cards = 1L)
pair <- policy_turn(pair)

writeLines(readable_log(pair$state, turn_number = 1L))
```

## Turn 2

```{r one_game_turn2}
pair$state <- begin_turn(pair$state)
pair <- draw_to_hand(pair, num_cards = 1L)
pair <- policy_turn(pair)

writeLines(readable_log(pair$state, turn_number = 2L))
```

## What the window closed on

```{r one_game_result}
result <- summarise_replicate(pair, decklist$decklist_id, seed_number,
                              bool_keep_trace = TRUE)

print(paste0("hit: ", result$bool_hit,
             "   turn achieved: ", result$jammer_turn))
writeLines(result$trace_vec)
```

That block is what a trace looks like. `end of turn 2` is recorded in full
(ADR 0007) because the board reached by turn 2 is a question worth asking later,
and the window stops there — there is no turn 3 anywhere in the simulator.

# Both cells, at scale

Going first and going second are different games and get different numbers. Each
cell is 1,000 replicates of the `clear` scenario.

```{r run_both_cells}
num_replicates <- 1000L

run_cell <- function(bool_going_first, scenario = "clear"){
  lapply(seq_len(num_replicates), function(i){
    one_pair <- play_replicate(decklist, card_df,
                               bool_going_first = bool_going_first,
                               scenario = scenario,
                               seed_number = i)
    summarise_replicate(one_pair, decklist$decklist_id, i)
  })
}

second_list <- run_cell(bool_going_first = FALSE)
first_list <- run_cell(bool_going_first = TRUE)

second_summary <- summarise_run(second_list)
first_summary <- summarise_run(first_list)

data.frame(cell = c("going second", "going first"),
           hit_rate = c(second_summary$hit_rate, first_summary$hit_rate),
           turn1 = c(second_summary$turn_tally_vec[["t1"]],
                     first_summary$turn_tally_vec[["t1"]]),
           turn2 = c(second_summary$turn_tally_vec[["t2"]],
                     first_summary$turn_tally_vec[["t2"]]),
           never = c(second_summary$turn_tally_vec[["never"]],
                     first_summary$turn_tally_vec[["never"]]),
           row.names = NULL)
```

Going second hits on turn 1 sometimes and going first never does — that is
Salvatore, and it is the only place its speed is visible, because the primary
outcome deliberately prices a turn-1 kill and a turn-2 conventional line the
same.

**Mulligans, reported beside the rate and never inside it** (ADR 0005):

```{r mulligan_metrics}
data.frame(cell = c("going second", "going first"),
           mulligan_rate = c(second_summary$mulligan_rate,
                             first_summary$mulligan_rate),
           mean_mulligans = c(second_summary$mean_mulligans,
                              first_summary$mean_mulligans),
           row.names = NULL)
```

## Which sub-goal actually fails

Counted as a **set**, so a miss appears in every row it belongs to. Rows sum to
more than the miss count on purpose: reporting only the first unmet sub-goal
made C look like it never failed, which is the exact opposite of the truth.

```{r unmet_tally}
data.frame(subgoal = names(SUBGOAL_VEC),
           meaning = as.character(SUBGOAL_VEC),
           going_second = as.integer(second_summary$unmet_tally_vec),
           going_first = as.integer(first_summary$unmet_tally_vec),
           row.names = NULL)
```

§1 of the decision tree claims **C is the sub-goal that actually fails** —
getting Bronzong *Active*, not drawing it. These counts are the first real test
of that claim.

## Hit rate by the Basic that led

§3's lead order is an untested default that you asked to settle from the logs.
This is the table that settles it — computed over every replicate, so unlike the
traces it may be read as a rate.

```{r lead_table}
lead_df <- second_summary$lead_hit_df
lead_df$lead <- lookup_card(card_df, lead_df$lead_card_id)$name
lead_df[order(-lead_df$hit_rate), c("lead", "num_replicates", "num_hit",
                                    "hit_rate")]
```

```{r lead_plot, fig.width=8, fig.height=4.5, fig.cap="Hit rate by opening lead, going second"}
plot_df <- lead_df[order(lead_df$hit_rate), ]
graphics::par(mar = c(4, 11, 2, 2))
graphics::barplot(plot_df$hit_rate, names.arg = plot_df$lead, horiz = TRUE,
                  las = 1, xlim = c(0, 1), col = "#4C72B0",
                  xlab = "hit rate (Evolution Jammer by turn 2)")
graphics::abline(v = second_summary$hit_rate, lty = 2, col = "#C44E52")
```

The dashed line is the cell's overall rate. Leading a Bronzor is worth far more
than any other lead, which is what §3 assumes.

**Two things in this table argue against §3 as written**, and both need a better
policy before they are worth acting on:

```{r lead_prose_values, include=FALSE}
rate_of <- function(name_str){
  paste0(format(round(100 * lead_df$hit_rate[lead_df$lead == name_str], 1),
                nsmall = 1), "%")
}
n_of <- function(name_str) lead_df$num_replicates[lead_df$lead == name_str]
```

- **§3's choice of Mega Kangaskhan ex looks right, but only once Run Errand is
  actually used** (`r rate_of("Mega Kangaskhan ex")`, n = `r n_of("Mega Kangaskhan ex")`).
  In the first draft of this policy it came out near the bottom at 40.9%, purely
  because the policy never used the Ability. Implementing a two-card draw moved
  one lead by 15 points — worth remembering before reading any of these numbers
  as facts about the *deck* rather than about the *policy*.
- **Buneary is §3's second choice and comes out second from last**
  (`r rate_of("Buneary")`, n = `r n_of("Buneary")`). Its whole case is Run
  Around, which §4.2 then makes a last resort because it spends the turn's
  Energy attachment. That may be the tree disagreeing with itself.
- **Latias ex is the best non-Bronzor lead here**
  (`r rate_of("Latias ex")`, n = `r n_of("Latias ex")`), and §3 says to lead it
  only when nothing else is available. Skyliner works from the Bench, so the
  tree's reasoning is sound — but leading it does get the free retreat online on
  turn 1 with no card spent. The sample is small; treat it as a question, not a
  result.

This is exactly the loop the trace machinery exists for: the tree makes a claim,
the run tests it, and the disagreement is where the next revision goes.

## Play motifs

Counted over every miss. These are the counts to act on: each one names a
pattern the policy produced, so a high count is a decision to change rather than
a card to add.

```{r motif_tally}
data.frame(motif = as.character(MOTIF_VEC),
           going_second = as.integer(second_summary$motif_tally_vec),
           going_first = as.integer(first_summary$motif_tally_vec),
           row.names = NULL)
```

# The trace file

The second output. A run writes a small sample of readable traces, **stratified
toward misses on purpose** (ADR 0006), because hits teach nothing about what to
change.

```{r write_traces}
sampler <- new_trace_sampler(max_miss = 8L, max_hit = 2L)
trace_list <- list()

for(i in seq_along(second_list)){
  take_list <- sampler_take(sampler, second_list[[i]]$bool_hit)
  sampler <- take_list$sampler
  if(!take_list$bool_keep) next

  one_pair <- play_replicate(decklist, card_df, bool_going_first = FALSE,
                             seed_number = i)
  trace_list[[length(trace_list) + 1]] <- summarise_replicate(
    one_pair, decklist$decklist_id, i, bool_keep_trace = TRUE)
}

trace_path <- "demo_traces.txt"
write_trace_file(trace_list, trace_path, second_summary, decklist = decklist)
```

> **Never compute a rate from a trace file.** The sample is deliberately not
> representative of the outcome distribution — it is roughly 80% misses by
> construction. The file leads with the true rates for exactly this reason, and
> the natural thing for a reader to do is count the traces anyway.

Two of the sampled traces:

```{r show_traces}
line_vec <- readLines(trace_path)
start_idx <- grep("^#[0-9]+/", line_vec)
writeLines(line_vec[start_idx[1]:(start_idx[3] - 1)])
```

Each trace carries two fields that exist to separate a **deck** problem from a
**decision** problem. `unmet=` names the sub-goals that were still open, and a
`!! PLAYABLE OUT unused` line means the card that would have fixed it was in
hand and was not played — that is a bug in `docs/03_decision_tree.md`, not in
the 60 cards.

# A question the demo raised for the tree

The narrated game above ends with **Salvatore in hand, unplayed, on turn 2** —
flagged by the trace as a playable out. That is the policy following §6, whose
priority 1 reads *"Turn 1 going second, and the kill is live. Never otherwise on
turn 1 — its exemption is wasted on turn 2."* Read literally, Salvatore is never
played on turn 2 at all.

But on turn 2 Salvatore still fetches Bronzong **and** evolves in one card, which
is a perfectly good turn-2 play when Hilda is not in hand. The wasted part is its
first-turn exemption, not the card. **Should §6 list Salvatore as a turn-2
Supporter, ranked against Hilda?** In this game it would have whiffed — Bronzong
was prized, which the player could not have known — and even the whiff would have
been worth having, since it reveals the prize.

# What the policy does not do yet

Stated plainly so your feedback lands on decisions rather than on known gaps.

- **It leaves the Supporter slot unspent far too often** — the single biggest
  motif count above. Every Supporter in the deck is implemented, but the policy
  reaches for Hilda first and often finds nothing else it judges worth playing.
  Note this count went *up* when the policy got better: a Hilda played to fetch
  nothing used to count as the slot being spent.
- **It never plays Pokégear 3.0**, which is the only Item that digs for a
  Supporter and is exactly what the going-first branch lacks.
- **It never takes the Cursed Blast escape** (§8). The machinery exists —
  `play_rare_candy()`, `use_cursed_blast()`, `knock_out()` — but the policy
  never sets up the Duskull-Active line that reaches it.
- **Retreat is only taken when free.** A paid retreat is never considered, even
  when the Energy spent would have been wasted anyway.
- **Boss's Orders is the one Supporter with no implementation**, deliberately:
  it moves the *opponent's* Pokémon, and no opponent is modelled.
- **The `item_lock` scenario is not shown here** — only `clear`. It is
  implemented and runs; it just is not in this document.

# What we glossed over

- **The opponent is a scenario, not a board.** No opposing decklist is modelled,
  which is also why the opening hand omits the bonus cards owed for an
  opponent's mulligans — biasing every rate slightly downward by an unknown
  amount.
- **Only one decklist appears here.** Part 6 is the registry that runs all six
  across both cells and both scenarios; this document runs one cell pair by hand.
- **The lead order and Ciphermaniac's timing are defaults, not rulings.** Both
  are written to be overturned by the tables above once the policy is good
  enough for those tables to mean something.

# References

- `docs/01_rules_standard.md` — the rules this engine implements
- `docs/03_decision_tree.md` — the tree this policy translates, §9 for the
  answers and corrections so far
- `docs/adr/` — the seven decisions that were expensive to make
- `CONTEXT.md` — what this project's words mean
