---
title: "How the Bronzong simulator works"
author: "Kevin Z. Lin"
date: "2026-08-29"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Overview

This is a walkthrough of the simulator as it stands, built to be argued with.
The policy it runs — `R/decision_claude.R` — is a **first draft translation of
`docs/03_decision_tree.md`**, written so the tree and the code can be read side
by side. It will make choices you disagree with. Those disagreements are the
output this document is for.

**The question the whole thing answers:** over many simulated games, how often
can the player attack with **Evolution Jammer on or before their own turn 2**?
One fixed bar, the same going first and going second (ADR 0004), reported
separately for the two coin-flip branches and never pooled (ADR 0002).

| Stage | What it does | Functions |
|---|---|---|
| Setup | Shuffle, mulligan, place the Active, set prizes | `setup_game()`, `policy_placement()` |
| Turns 1–2 | Draw, then play the decision tree | `policy_turn()`, `play_replicate()` |
| Measure | One replicate → hit / miss, turn achieved, diagnosis | `summarise_replicate()` |
| Aggregate | Many replicates → one rate per cell | `summarise_run()` |
| Diagnose | A stratified sample of readable traces | `write_trace_file()` |

Two things the simulator is careful about, because both are easy to get silently
wrong:

- **The policy may not see hidden information** (ADR 0003). It reads the hand,
  the board and the discard, plus a belief state that only learns what a real
  player would learn. A search that fails because the card was prized is
  information the player *earns*.
- **Mulligans are never a miss** (ADR 0005). A hand with no Basic is redrawn and
  the redraw is played; the cost is reported beside the rate, never folded into
  it.

```{r libraries, message=FALSE}
for(one_file in list.files("../R", pattern = "[.]R$", full.names = TRUE)){
  source(one_file)
}

card_df <- build_card_database()
decklist <- read_decklist("../decklists/decklist2.txt", card_df)
```

# The deck

`decklist2`, chosen because it runs **both** Salvatore and Latias ex — so the
turn-1 kill line and the free-retreat rung of the §4.3 ladder both get
exercised.

```{r show_decklist}
count_vec <- decklist$count_vec
data.frame(count = as.integer(count_vec),
           card = lookup_card(card_df, names(count_vec))$name,
           row.names = NULL)
```

The four sub-goals that must all hold at the attack step, from §1 of the
decision tree:

| | Sub-goal | Satisfied by |
|---|---|---|
| **A** | A Bronzor is in play | setup, Poké Pad, Ultra Ball, Brock's Scouting, benching one |
| **B** | Bronzong is on it | Hilda, Salvatore, or drawn and evolved |
| **C** | Bronzong is **Active** | led at setup, free retreat, Switch, Surfer, Run Around, Cursed Blast |
| **D** | A `[P]` source is attached | basic Psychic or Telepathic Psychic Energy |

# One game, played slowly

The most useful part of this document. A single seed, going second, played turn
by turn — every decision the policy takes is logged as it happens.

```{r one_game_setup}
seed_number <- 8L

pair <- setup_game(decklist, card_df,
                   bool_going_first = FALSE,
                   placement_fn = policy_placement,
                   seed_number = seed_number)

print(paste0("opening hand: ",
             paste0(lookup_card(card_df, pair$state$hand_vec)$name,
                    collapse = ", ")))
print(paste0("led: ", lookup_card(card_df, top_card(pair$state$active))$name))
print(paste0("bench: ", length(pair$state$bench_list), " Pokemon"))
```

Note the empty Bench. §3 says to place the Active and stop: a benched Basic can
never be un-benched, and the slots are wanted for Latias ex.

## Turn 1

```{r one_game_turn1}
pair$state <- begin_turn(pair$state)
pair <- draw_to_hand(pair, num_cards = 1L)
pair <- policy_turn(pair)

writeLines(readable_log(pair$state, turn_number = 1L))
```

## Turn 2

```{r one_game_turn2}
pair$state <- begin_turn(pair$state)
pair <- draw_to_hand(pair, num_cards = 1L)
pair <- policy_turn(pair)

writeLines(readable_log(pair$state, turn_number = 2L))
```

## What the window closed on

```{r one_game_result}
result <- summarise_replicate(pair, decklist$decklist_id, seed_number,
                              bool_keep_trace = TRUE)

print(paste0("hit: ", result$bool_hit,
             "   turn achieved: ", result$jammer_turn))
writeLines(result$trace_vec)
```

That block is what a trace looks like. `end of turn 2` is recorded in full
(ADR 0007) because the board reached by turn 2 is a question worth asking later,
and the window stops there — there is no turn 3 anywhere in the simulator.

# Both cells, at scale

Going first and going second are different games and get different numbers. Each
cell is 1,000 replicates of the `clear` scenario.

```{r run_both_cells}
num_replicates <- 1000L

run_cell <- function(bool_going_first, scenario = "clear"){
  lapply(seq_len(num_replicates), function(i){
    one_pair <- play_replicate(decklist, card_df,
                               bool_going_first = bool_going_first,
                               scenario = scenario,
                               seed_number = i)
    summarise_replicate(one_pair, decklist$decklist_id, i)
  })
}

second_list <- run_cell(bool_going_first = FALSE)
first_list <- run_cell(bool_going_first = TRUE)

second_summary <- summarise_run(second_list)
first_summary <- summarise_run(first_list)

data.frame(cell = c("going second", "going first"),
           hit_rate = c(second_summary$hit_rate, first_summary$hit_rate),
           turn1 = c(second_summary$turn_tally_vec[["t1"]],
                     first_summary$turn_tally_vec[["t1"]]),
           turn2 = c(second_summary$turn_tally_vec[["t2"]],
                     first_summary$turn_tally_vec[["t2"]]),
           never = c(second_summary$turn_tally_vec[["never"]],
                     first_summary$turn_tally_vec[["never"]]),
           row.names = NULL)
```

Going second hits on turn 1 sometimes and going first never does — that is
Salvatore, and it is the only place its speed is visible, because the primary
outcome deliberately prices a turn-1 kill and a turn-2 conventional line the
same.

**Mulligans, reported beside the rate and never inside it** (ADR 0005):

```{r mulligan_metrics}
data.frame(cell = c("going second", "going first"),
           mulligan_rate = c(second_summary$mulligan_rate,
                             first_summary$mulligan_rate),
           mean_mulligans = c(second_summary$mean_mulligans,
                              first_summary$mean_mulligans),
           row.names = NULL)
```

## Which sub-goal actually fails

Counted as a **set**, so a miss appears in every row it belongs to. Rows sum to
more than the miss count on purpose: reporting only the first unmet sub-goal
made C look like it never failed, which is the exact opposite of the truth.

```{r unmet_tally}
data.frame(subgoal = names(SUBGOAL_VEC),
           meaning = as.character(SUBGOAL_VEC),
           going_second = as.integer(second_summary$unmet_tally_vec),
           going_first = as.integer(first_summary$unmet_tally_vec),
           row.names = NULL)
```

§1 of the decision tree claims **C is the sub-goal that actually fails** —
getting Bronzong *Active*, not drawing it. These counts are the first real test
of that claim.

## Hit rate by the Basic that led

**This table cannot settle §3's lead order, and it is worth saying why before
reading it.** It groups every replicate by the Basic that led — but the hand that
contains a Mega Kangaskhan ex is not the hand that contains a Duskull, so it
compares leads *across different hands* and reports a property of the hands as
though it were a property of the order. The lead is not randomised here; it is
chosen, and it is chosen by the very rule under test.

The order is settled instead by **varying it and measuring the cell rate**, which
is what `scripts/tune_lead_order_claude.R` does and what ADR 0008 records. Read
the table below as a description of what the policy did.

```{r lead_table}
lead_df <- second_summary$lead_hit_df
lead_df$lead <- lookup_card(card_df, lead_df$lead_card_id)$name
lead_df[order(-lead_df$hit_rate), c("lead", "num_replicates", "num_hit",
                                    "hit_rate")]
```

```{r lead_plot, fig.width=8, fig.height=4.5, fig.cap="Hit rate by opening lead, going second"}
plot_df <- lead_df[order(lead_df$hit_rate), ]
graphics::par(mar = c(4, 11, 2, 2))
graphics::barplot(plot_df$hit_rate, names.arg = plot_df$lead, horiz = TRUE,
                  las = 1, xlim = c(0, 1), col = "#4C72B0",
                  xlab = "hit rate (Evolution Jammer by turn 2)")
graphics::abline(v = second_summary$hit_rate, lty = 2, col = "#C44E52")
```

The dashed line is the cell's overall rate. Leading a Bronzor is worth far more
than any other lead, which is what §3 assumes.

**Both of the things this table pointed at turned out to be real**, and the
proper experiment then confirmed one of them and rejected the other:

```{r lead_prose_values, include=FALSE}
rate_of <- function(name_str){
  paste0(format(round(100 * lead_df$hit_rate[lead_df$lead == name_str], 1),
                nsmall = 1), "%")
}
n_of <- function(name_str) lead_df$num_replicates[lead_df$lead == name_str]
```

- **§3's choice of Mega Kangaskhan ex looks right, but only once Run Errand is
  actually used** (`r rate_of("Mega Kangaskhan ex")`, n = `r n_of("Mega Kangaskhan ex")`).
  In the first draft of this policy it came out near the bottom at 40.9%, purely
  because the policy never used the Ability. Implementing a two-card draw moved
  one lead by 15 points — worth remembering before reading any of these numbers
  as facts about the *deck* rather than about the *policy*.
- **Buneary was §3's second choice going second and is now last**
  (`r rate_of("Buneary")`, n = `r n_of("Buneary")`). Its whole case is Run
  Around, which §4.2 then makes a last resort because it spends the turn's
  Energy attachment and strands it on the Bench. A lead whose one virtue the
  rest of the tree declines to use is not a virtue — the tree was disagreeing
  with itself, and the search resolved it.
- **Latias ex now leads going second** (`r rate_of("Latias ex")`,
  n = `r n_of("Latias ex")`), which §3 previously ranked last. Skyliner works
  from the Bench, so leading it looks like a waste; what leading it actually buys
  is the free retreat online on turn 1 at no card and no Bench slot.
- **And the size matters more than the ranking. The lead order is worth about a
  point.** Across every candidate order the going-second rate spans well under
  two points, against the 23 the Supporter rules moved. Going *first*, three
  different orders land within 0.22 points of each other, so §3's going-first
  order was left exactly as it was rather than replaced by the largest of twenty
  noisy numbers. `results/lead_order_tuning.md` has both searches in full.

This is exactly the loop the trace machinery exists for: the tree makes a claim,
the run tests it, and the disagreement is where the next revision goes.

## Play motifs

Counted over every miss. These are the counts to act on: each one names a
pattern the policy produced, so a high count is a decision to change rather than
a card to add.

```{r motif_tally}
data.frame(motif = as.character(MOTIF_VEC),
           going_second = as.integer(second_summary$motif_tally_vec),
           going_first = as.integer(first_summary$motif_tally_vec),
           row.names = NULL)
```

**"A turn ended with the Supporter slot unspent" is still the largest count, and
it is no longer a defect.** §6 priority 8 now plays a Supporter whenever one can
do anything, so the remaining count is hands that had none to play. Going second,
393 replicates in 1,000 end turn 2 with the slot unspent, and only **80** of those
still hold a Supporter at all — 28 a Salvatore with no Bronzor in play to put a
Bronzong onto, which is not a legal declaration, and 56 a Ciphermaniac's, whose
stack the window can never draw. **Hilda and Lillie's are never stranded.** The
motif is worth keeping because it would catch a regression, but the number to
watch is 80, not 393.

# The trace file

The second output. A run writes a small sample of readable traces, **stratified
toward misses on purpose** (ADR 0006), because hits teach nothing about what to
change.

```{r write_traces}
sampler <- new_trace_sampler(max_miss = 8L, max_hit = 2L)
trace_list <- list()

for(i in seq_along(second_list)){
  take_list <- sampler_take(sampler, second_list[[i]]$bool_hit)
  sampler <- take_list$sampler
  if(!take_list$bool_keep) next

  one_pair <- play_replicate(decklist, card_df, bool_going_first = FALSE,
                             seed_number = i)
  trace_list[[length(trace_list) + 1]] <- summarise_replicate(
    one_pair, decklist$decklist_id, i, bool_keep_trace = TRUE)
}

trace_path <- "demo_traces.txt"
write_trace_file(trace_list, trace_path, second_summary, decklist = decklist)
```

> **Never compute a rate from a trace file.** The sample is deliberately not
> representative of the outcome distribution — it is roughly 80% misses by
> construction. The file leads with the true rates for exactly this reason, and
> the natural thing for a reader to do is count the traces anyway.

Two of the sampled traces:

```{r show_traces}
line_vec <- readLines(trace_path)
start_idx <- grep("^#[0-9]+/", line_vec)
writeLines(line_vec[start_idx[1]:(start_idx[3] - 1)])
```

Each trace carries two fields that exist to separate a **deck** problem from a
**decision** problem. `unmet=` names the sub-goals that were still open, and a
`!! PLAYABLE OUT unused` line means the card that would have fixed it was in
hand and was not played — that is a bug in `docs/03_decision_tree.md`, not in
the 60 cards.

# Where the rate came from

The policy is aligned to the decision documents as of 2026-08-30, after two
rounds of your answers in `docs/03b_scenarios.md`.

**The first round moved the rate 23 points.** Going second **52.6% → 75.8%**,
going first **53.4% → 64.1%**. Almost all of it was one idea stated twice: §6
priority 8 plays a Supporter every turn one is legal (worth **15.2** going
second), and §7 step 6 then assembles the turn *again* over whatever it drew
(another **5.0** going second, **6.7** going first). Without that second pass the
fallback is worth nothing at all on turn 2, because ADR 0007 closes the window
before the eight cards can be played.

**The second round moved it 1.7.** Going second **75.8% → 77.5%**. That is the
shape of a policy that is mostly right: seven of your twelve answers confirmed
the tree, and the changes are refinements rather than structural gaps. The
per-change table is in `results/change_attribution.md`, and the number to read
first is its noise floor — at 1,000 paired replicates the standard error on one
of those differences is 0.2 to 0.4 points, so most of the rows are honestly zero.

**Two of them are zero on purpose.** S-18 (a spare `[P]` onto a second Bronzong)
and S-19 (Rare Candy once Bronzong is settled) change the **board the window
closes on**, not whether the attack happens. A 0.0 there is the answer rather
than a disappointment, and it is why the end-of-turn-2 snapshot is recorded in
full.

**The one position the tree could win and did not was S-20**, worth 0.3 points
and worth more than that as a lesson: with a Duskull Active and the line stranded
on the Bench, the Cursed Blast escape was one Dusknoir away — and Dusknoir is an
Evolution Pokémon, so Hilda could always have fetched it. The want-list simply
had no entry for it.

# What the policy does not do yet

Stated plainly so your feedback lands on decisions rather than on known gaps.

- **Three cards in decklist7 and decklist8 are transcribed and unimplemented**:
  Dunsparce's Trading Places (a `[C]` attack that switches, so Run Around's
  shape), Dudunsparce's Run Away Draw, and basic Darkness Energy as Run Around's
  cheapest payment. Each biases decklist8's rate *downward*, so its number is a
  lower bound. `docs/03a_card_playbook.md` → *Specified and not implemented*.
- **It never plays Pokégear 3.0**, which is the only Item that digs for a
  Supporter and is exactly what the going-first branch lacks.
- **Retreat is only taken when free.** A paid retreat is never considered, even
  when the Energy spent would have been wasted anyway.
- **Boss's Orders is the one Supporter with no implementation**, deliberately:
  it moves the *opponent's* Pokémon, and no opponent is modelled.
- **Risky Ruins' damage is modelled as not arising**, because §4.2 step 7
  declines to play it. That is correct only while the decline holds; a rule that
  ever plays it has to implement the damage in the same change.

# What we glossed over

- **The opponent is a scenario, not a board.** No opposing decklist is modelled,
  which is also why the opening hand omits the bonus cards owed for an
  opponent's mulligans — biasing every rate slightly downward by an unknown
  amount.
- **Only one decklist appears here.** `results/registry.md` is the registry that
  runs all eight across all three cells; this document runs one cell pair by hand
  because its job is to show the *machinery*, not to rank the decks.
- **Blissey ex cannot be played in either list that runs it.** It evolves from
  Chansey and no decklist runs one. That is a decklist question rather than a
  simulator one, and it is the kind of thing the registry cannot tell you.

# References

- `docs/01_rules_standard.md` — the rules this engine implements
- `docs/03_decision_tree.md` — the tree this policy translates, §9 for the
  questions it still states as defaults rather than rulings
- `docs/adr/` — the seven decisions that were expensive to make
- `CONTEXT.md` — what this project's words mean
